Incentivizing Exploration with Unbiased History

### 2018-10-27 2level.tex
3level

\begin{remark}
  If the gap parameter $\Delta$ is known to the principal, we can set
  $T_1 = \max\left( \frac{4\GdT^2}{\GdP^2}\log(T), \frac{36 }{\Delta^2
      \cdot p_G} \log(T) \right)$, and achieves an expected regret of
  $O(\log(T) \cdot \Delta^{-2})$.
\end{remark}

>>>>

\begin{remark}
  If the gap parameter $\Delta$ is known to the principal, we can set
  $$T_1 = \max\left( \frac{4\GdT^2}{\GdP^2}\log(T), \frac{36 }{\Delta^2
      \cdot p_G} \log(T) \right)$$, and achieves an expected regret of
  $O(\log(T) \cdot \Delta^{-2})$.
\end{remark}

### 2018-10-27 3level.tex
3level

\section{3-level Adaptive Policy}
\label{sec:3level}
\swcomment{rewrote below}
Now we demonstrate how to design a 3-level policy to incentivize
adaptive exploration and improve the $O(T^{2/3})$ regret rate to
$\tilde O(T^{4/7})$. In \Cref{sec:llevel}, we will show how to extend
this idea to achieve a regret rate of $\tilde O(T^{1/2})$. For
simplicity, we will focus on the setting with 2 arms.

@@

\begin{lemma}[Concentration of first-level number of pulls.]
  Let $W_1$ be the event that for all groups $s\in [S]$ and arms
  $a\in \{1, 2\}$, the number of arm $a$ pulls in the $s$-th
  first-level group is in the range of
  $$
  \left[q_a T_1- \GdT \sqrt{T_1\log(T)}, q_a T_1 + \GdT \sqrt{T_1\log(T)}\right],
  $$
  where $q_a$ is the expected number of arm $a$ pulls in a $\ALGG$ run
  of length $T_G$. Then $\Pr[W_1] \geq 1- \frac{4S}{T^2}$.
\end{lemma}

\begin{proof}
  For the $s$-th first-level group, define $W_1^{a,s}$ to be the event
  that the number of arm $a$ pulls in the $s$-th first-level group is
  between $q_a T_1- \GdT \sqrt{T_1\log(T)}$ and
  $q_a T_1 + \GdT \sqrt{T_1\log(T)}$. By Lemma~\ref{lem:t1runs}
\[
\Pr[W_1^{a,s}] \geq 1-2\exp(-2\log(T)) \geq 1-2/T^2.
\]
Let $W_1$  be the intersection of all these events (i.e.
$W_1 = \bigcap_{a,s}W_1^{a,s}$). By union bound, we have
\[
\Pr[W_1] \geq 1- \frac{4S}{T^2}.
\]
\swcomment{let's put this in appendix.}
\end{proof}

@@

\begin{lemma}[Concetration of empirical means in the first level]
  For any $t_1, t_2\in [T]$ such that $t_1 < t_2$, $s\in [S]$, and
  $a\in \{1,2\}$, let $W_2^{s,a,t_1,t_2}$ be the event that the mean
  among the cells indexed by $t_1, (t_1+1), \ldots, t_2$ in the tape
  $\cT^1_{a, s}$ is at most $\sqrt{\frac{2\log(T)}{t_2-t_1+1}}$ away
  from $\mu_a$.  Let $W_2$ be the intersection of all these events
  (i.e.  $W_2 = \bigcap_{a,s,t_1,t_2} W_2^{s,a,t_1,t_2}$). Then
  \[
    \Pr[W_2] \geq 1- \frac{4S}{T^2}.
  \]
\end{lemma}

\OMIT{
\begin{proof}
  By Chernoff bound,
\[
\Pr[W_2^{s,a,t_1,t_2}] \geq 1 - 2\exp(-4\log(T)) \geq 1- 2/T^4.
\]
By union bound, we have
\[
\Pr[W_2] \geq 1- \frac{4S}{T^2}.
\]
\end{proof}
}

@@

\begin{lemma}[High and low deviations for the arms]
  For any group $s\in [S]$, any arm $a$, let $\tilde\mu_{a,s}$ be the
  empirical mean of the first $q_a T_1$ cells in tape $\cT^1_{a, s}$.
  Let $W_4^{s,a,high}$ be the event
  $\tilde\mu_{a, s} \geq \mu_a + 1/\sqrt{q_a T_1}$ and let
  $W_4^{s,a,low}$ be the event that
  $\tilde\mu_{a, s} \leq \mu_a - 1/\sqrt{q_a T_1}$.  Let $W_4$ be the
  event that for every $a\in \{1, 2\}$, there exists a group
  $s_a\in [S]$ in the first level such that both $W_4^{s_a,a,high}$
  and $W_4^{s_a,3-a,low}$ occur. Then
  \[
    \Pr[W_4]\geq 1 -2 /T.
  \]
\end{lemma}

\begin{proof}
By Berry-Essen Theorem and
  $\mu_a \in [1/3,2/3]$, we have for any $a$,
\[
\Pr[W_4^{s,a,high}] \geq (1-\Phi(1/2)) - \frac{5}{\sqrt{q_aT_1}} > 1/4.
\]
The last inequality follows when $T$ is larger than some constant.
Similarly we also have 
\[
\Pr[W_4^{s,a,low}] > 1/4.
\]
Since $W_4^{s,a,high}$ is independent with $W_4^{s,3-a,low}$, we have
\[
\Pr[W_4^{s,a,high} \cap W_4^{s,3-a,low}] =\Pr[W_4^{s,a,high}] \cdot  \Pr[W_4^{s,3-a,low}]>(1/4)^2 = 1/16.
\]
Notice that $(W_4^{s,a,high} \cap W_4^{s,3-a,low})$ are independent
across different $s$'s. By union bound, we have
\[
\Pr[W_4] \geq 1- 2(1-1/16)^S \geq 1 -2 /T.
\]
\end{proof}

Lastly, we will condition on the event that the empirical means of
both arms to be concentrated around their true means in any prefix of
their pulls.
% We will again state the event using a hypothetical reward tape
% $\cT^2_a$ for each arm $a$: the $j$-th time arm $a$ is chosen in the
% two levels, its reward is taken from the $j$-th cell in $\cT^2_a$.

\begin{lemma}
  With probability at least $1 - \frac{4}{T^3}$, the following event
  $W_3$ holds: for all $a\in \{1, 2\}$ and $t\in [T]$, the empirical
  means of the first $t$ arm $a$ pulls is at most
  $\sqrt{\frac{2\log(T)}{t}}$ away from $\mu_a$.
\end{lemma}

\swcomment{might want to swap out $W_3$ and $W_4$}

\OMIT{\begin{proof}
By Chernoff bound,
\[
\Pr[W_3^{a,t}] \geq 1 - 2\exp(-4\log(T)) \geq 1- 2/T^4.
\]
By union bound, we have
\[
\Pr[W_3] \geq 1- \frac{4}{T^3}.
\]
\end{proof}}

\subsection{Case Analysis}

\begin{proof}
Wlog we assume $\mu_1 \geq \mu_2$ as the recommendation policy is symmetric to both arms. We do a case analysis based on $\mu_1-\mu_2$.

Before we start with the case analysis, we first define several clean events and show that the intersection of them happens with high probability. 
\OMIT{
\begin{itemize}

@@

Consider the tapes defined in the second bullet again. For the $s$-th first-level group and arm $a$, define $W_4^{s,a,high}$  to be the event that first $q_a T_1$ pulls of arm $a$ in the corresponding tape has empirical mean at least $\mu_a + 1/\sqrt{q_a T_1}$ and define  $W_4^{s,a,low}$  to be the event that first $q_a T_1$ pulls of arm $a$ in the corresponding tape has empirical mean at most $\mu_a - 1/\sqrt{q_a T_1}$. By Berry-Essen Theorem and $\mu_a \in [1/3,2/3]$, we have
\[
\Pr[W_4^{s,a,high}] \geq (1-\Phi(1/2)) - \frac{5}{\sqrt{q_aT_1}} > 1/4.
\]
The last inequality follows when $T$ is larger than some constant.
Similarly we also have 
\[
\Pr[W_4^{s,a,low}] > 1/4.
\]
Since $W_4^{s,a,high}$ is independent with $W_4^{s,3-a,low}$, we have
\[
\Pr[W_4^{s,a,high} \cap W_4^{s,3-a,low}] =\Pr[W_4^{s,a,high}] \cdot  \Pr[W_4^{s,3-a,low}]>(1/4)^2 = 1/16.
\]
Now define $W_4$ as $\bigcap_a \bigcup_s (W_4^{s,a,high} \cap W_4^{s,3-a,low})$. Notice that $(W_4^{s,a,high} \cap W_4^{s,3-a,low})$ are independent across different $s$'s. By union bound, we have
\[
\Pr[W_4] \geq 1- 2(1-1/16)^S \geq 1 -2 /T.
\]
\end{itemize}}

@@

Now we assume the intersection of clean events happens and we summarize what these clean events imply.

@@

\begin{itemize}
\item $\mu_1 - \mu_2 \geq 2\left(\sqrt{\frac{4\log(T)}{q_1T_1}} 
+ \sqrt{\frac{4\log(T)}{q_2T_1}}\right)$. In this case, we want to show that agents in the second and the third levels all pull arm 1.

@@

Now consider the agents in the third level group. Recall $\bar{\mu}_a$ is the empirical mean of arm $a$ in the history they see. We have
\[
\bar{\mu}_1 - \bar{\mu}_2 \geq \mu_1 -\mu_2 - \sqrt{\frac{4\log(T)}{Sq_1T_1}} - \sqrt{\frac{4\log(T)}{Sq_2T_1}} \geq  \sqrt{\frac{4\log(T)}{q_1T_1}} 
+ \sqrt{\frac{4\log(T)}{q_2T_1}}.
\]
Similarly as above, by Assumption \ref{ass:embehave}, we know $\hat{\mu}_1^t - \hat{\mu}_2^t > 0$ for any agent $t$ in the third level. So we know agents in the third-level group will all pull arm 1. Therefore the expected regret is at most $S T_G T_1 = O(T^{4/7} \log^{6/7}(T))$.

@@

\item $\mu_1 - \mu_2 \leq 3\sqrt{\frac{2\log(T)}{T_2}}$. This is the easy case. Even always pulling the sub-optimal arm (i.e. arm 2) gives regret at most $T \cdot (\mu_1-\mu_2) = O(T^{4/7} \log^{6/7}(T))$. 
\end{itemize}

>>>>

\section{Adaptive Policy with Three Levels}
\label{sec:3level}
\swcomment{rewrote below}
Now we demonstrate how to design a 3-level policy to incentivize
adaptive exploration and improve the $O(T^{2/3})$ regret rate to
$\tilde O(T^{4/7})$. In \Cref{sec:llevel}, we will show how to extend
this idea to achieve a regret rate of $\tilde O(T^{1/2})$. For
simplicity, we will focus on the setting with 2 arms.

@@

\begin{lemma}[Concentration of first-level number of pulls.]\label{3levelw1}
  Let $W_1$ be the event that for all groups $s\in [S]$ and arms
  $a\in \{1, 2\}$, the number of arm $a$ pulls in the $s$-th
  first-level group is in the range of
  $$
  \left[q_a T_1- \GdT \sqrt{T_1\log(T)}, q_a T_1 + \GdT \sqrt{T_1\log(T)}\right],
  $$
  where $q_a$ is the expected number of arm $a$ pulls in a $\ALGG$ run
  of length $T_G$. Then $\Pr[W_1] \geq 1- \frac{4S}{T^2}$.
\end{lemma}

@@

\begin{lemma}[Concentration of empirical means in the first level]\label{3levelw2}
  For any $t_1, t_2\in [T]$ such that $t_1 < t_2$, $s\in [S]$, and
  $a\in \{1,2\}$, let $W_2^{s,a,t_1,t_2}$ be the event that the mean
  among the cells indexed by $t_1, (t_1+1), \ldots, t_2$ in the tape
  $\cT^1_{a, s}$ is at most $\sqrt{\frac{2\log(T)}{t_2-t_1+1}}$ away
  from $\mu_a$.  Let $W_2$ be the intersection of all these events
  (i.e.  $W_2 = \bigcap_{a,s,t_1,t_2} W_2^{s,a,t_1,t_2}$). Then
  \[
    \Pr[W_2] \geq 1- \frac{4S}{T^2}.
  \]
\end{lemma}

@@

\begin{lemma}[High and low deviations for the arms]\label{3levelw4}
  For any group $s\in [S]$, any arm $a$, let $\tilde\mu_{a,s}$ be the
  empirical mean of the first $q_a T_1$ cells in tape $\cT^1_{a, s}$.
  Let $W_4^{s,a,high}$ be the event
  $\tilde\mu_{a, s} \geq \mu_a + 1/\sqrt{q_a T_1}$ and let
  $W_4^{s,a,low}$ be the event that
  $\tilde\mu_{a, s} \leq \mu_a - 1/\sqrt{q_a T_1}$.  Let $W_4$ be the
  event that for every $a\in \{1, 2\}$, there exists a group
  $s_a\in [S]$ in the first level such that both $W_4^{s_a,a,high}$
  and $W_4^{s_a,3-a,low}$ occur. Then
  \[
    \Pr[W_4]\geq 1 -2 /T.
  \]
\end{lemma}

Lastly, we will condition on the event that the empirical means of
both arms to be concentrated around their true means in any prefix of
their pulls. This guarantees that the policy obtains an accurate
estimate of rewards for both arms after aggregating all the data in
the first two levels.
% We will again state the event using a hypothetical reward tape
% $\cT^2_a$ for each arm $a$: the $j$-th time arm $a$ is chosen in the
% two levels, its reward is taken from the $j$-th cell in $\cT^2_a$.

\swcomment{might want to swap out $W_3$ and $W_4$}
\begin{lemma}[Concentration of empirical means in the first two
  levels]\label{3levelw3}
  With probability at least $1 - \frac{4}{T^3}$, the following event
  $W_3$ holds: for all $a\in \{1, 2\}$ and $m \in [N_{T, a}]$, the
  empirical means of the first $t$ arm $a$ pulls is at most
  $\sqrt{\frac{2\log(T)}{m}}$ away from $\mu_a$, where $N_{T, a}$ is
  the total number of arm $a$ pulls by the end of $T$ rounds.
\end{lemma}

Let $W = \bigcap_{i=1}^4 W_i$ be the intersection of all 4 clean
events.  By union bound, $W$ occuers with probability $1-O(1/T)$. Note
that expected regret conditioned on $W$ not occuring is at most
$O(1/T) \cdot T = O(1)$, so it suffices to bound the regret under $W$.

\subsection{Case Analysis under Clean Events}

We now proceed to bound the regret conditioned on all four clean
events defined above occur. Without loss of generality, we assume
$\mu_1 \geq \mu_2$ as the recommendation policy is symmetric to both
arms. We break down the regret analysis into 4 cases on the scale of
the gap parameter $\Delta = (\mu_1-\mu_2)$.

We will start with the simplest case where the gap is smaller than
$\tilde O(\sqrt{1/T_2}) = \tilde O(1/T^{3/7})$, then the regret is
trivially bounded by $\tilde O(T^{4/7})$.

\begin{claim}[Small gap case]
  Suppose $\Delta \leq 3\sqrt{\frac{2\log(T)}{T_2}}$,
  $\reg(T)\leq O(T^{4/7} \log^{6/7}(T))$.
\end{claim}

Another simple case is when $\Delta$ is sufficiently large, so that
the data collected in the first level can distinguish the two
arms. The proof essentially follows from the clean event $W_1$ that
both arms are explored sufficient many times in level 1, and the
empirical rewards in the first two levels concentrate around their
true means.

\begin{lemma}[Large gap case]\label{3levelbigcase}
  Suppose
  $\Delta \geq 2\left(\sqrt{\frac{4\log(T)}{q_1T_1}} +
    \sqrt{\frac{4\log(T)}{q_2T_1}}\right)$, then all agents in the
  second and third levels pull arm 1.
\end{lemma}

The analysis becomes more challenging when the gap is neither too
small to have negligible effect on regret nor too big to make the arms
distinguishable after the first level.

\begin{proof}
\OMIT{Wlog we assume $\mu_1 \geq \mu_2$ as the recommendation policy is symmetric to both arms. We do a case analysis based on $\mu_1-\mu_2$.

Before we start with the case analysis, we first define several clean events and show that the intersection of them happens with high probability.

\begin{itemize}

@@

Consider the tapes defined in the second bullet again. For the $s$-th first-level group and arm $a$, define $W_4^{s,a,high}$  to be the event that first $q_a T_1$ pulls of arm $a$ in the corresponding tape has empirical mean at least $\mu_a + 1/\sqrt{q_a T_1}$ and define  $W_4^{s,a,low}$  to be the event that first $q_a T_1$ pulls of arm $a$ in the corresponding tape has empirical mean at most $\mu_a - 1/\sqrt{q_a T_1}$. By Berry-Essen Theorem and $\mu_a \in [1/3,2/3]$, we have
\[
\Pr[W_4^{s,a,high}] \geq (1-\Phi(1/2)) - \frac{5}{\sqrt{q_aT_1}} > 1/4.
\]
The last inequality follows when $T$ is larger than some constant.
Similarly we also have 
\[
\Pr[W_4^{s,a,low}] > 1/4.
\]
Since $W_4^{s,a,high}$ is independent with $W_4^{s,3-a,low}$, we have
\[
\Pr[W_4^{s,a,high} \cap W_4^{s,3-a,low}] =\Pr[W_4^{s,a,high}] \cdot  \Pr[W_4^{s,3-a,low}]>(1/4)^2 = 1/16.
\]
Now define $W_4$ as $\bigcap_a \bigcup_s (W_4^{s,a,high} \cap W_4^{s,3-a,low})$. Notice that $(W_4^{s,a,high} \cap W_4^{s,3-a,low})$ are independent across different $s$'s. By union bound, we have
\[
\Pr[W_4] \geq 1- 2(1-1/16)^S \geq 1 -2 /T.
\]
\end{itemize}

@@

Now we assume the intersection of clean events happens and we summarize what these clean events imply.}

@@

\begin{itemize}
\OMIT{
\item $\mu_1 - \mu_2 \geq 2\left(\sqrt{\frac{4\log(T)}{q_1T_1}} 
+ \sqrt{\frac{4\log(T)}{q_2T_1}}\right)$. In this case, we want to show that agents in the second and the third levels all pull arm 1.

@@

Now consider the agents in the third level group. Recall $\bar{\mu}_a$ is the empirical mean of arm $a$ in the history they see. We have
\[
\bar{\mu}_1 - \bar{\mu}_2 \geq \mu_1 -\mu_2 - \sqrt{\frac{4\log(T)}{Sq_1T_1}} - \sqrt{\frac{4\log(T)}{Sq_2T_1}} \geq  \sqrt{\frac{4\log(T)}{q_1T_1}} 
+ \sqrt{\frac{4\log(T)}{q_2T_1}}.
\]
Similarly as above, by Assumption \ref{ass:embehave}, we know $\hat{\mu}_1^t - \hat{\mu}_2^t > 0$ for any agent $t$ in the third level. So we know agents in the third-level group will all pull arm 1. Therefore the expected regret is at most $S T_G T_1 = O(T^{4/7} \log^{6/7}(T))$. 
}

@@

\OMIT{
\item $\mu_1 - \mu_2 \leq 3\sqrt{\frac{2\log(T)}{T_2}}$. This is the easy case. Even always pulling the sub-optimal arm (i.e. arm 2) gives regret at most $T \cdot (\mu_1-\mu_2) = O(T^{4/7} \log^{6/7}(T))$. 
}
\end{itemize}
